# HW #3

## Formatting Guidance ([SAMPLE FILE](https://drive.google.com/file/d/1t7YkI9z3lFiUyRwKfXYCgg6L_oUxSuap/view?usp=sharing))

-   Submission:
    -   You should submit two files for each lab– a .Rmd and a .html.
    -   You should not submit any other files.
    -   The files should be saved using the following convention: LASTNAME_HW_3.Rmd
-   Code:
    -   All R code must appear inside code chunks. Do not place any code in the body text.
    -   All commands **MUST** have a descriptive comment
    -   Code chunks should not display messages or warnings in the knitted .html. If messages or warnings are printed, you need to include `message = F, warning = F` in the r header.
    -   Each code chunk should perform one logical set of tasks/analyses (i.e. one code chunk could include all of your data manipulation, another code chunk could contain all of your data visualizations)
    -   Do not print large datasets or unnecessary intermediate outputs
-   Organization
    -   Code chunks should be logically organized under clear, informative section headers
    -   Written responses should appear directly below the code chunk that produces the the relevant output

## Homework Overview

In this homework you will practice executing both spatial and non-spatial wrangling to feature engineer three analytical datasets.

## Rubric

| Specification | Points |
|----|----|
| Lab is submitted on Canvas and follows general formatting guidelines. Minor formatting issues may be present but do not interfere with readability or interpretation. | 5 |
| **RQ1:** Processing directions are followed and map clearly represents final analytical dataset. | 5 |
| **RQ2:** Processing directions are followed and map clearly represents final analytical dataset. | 5 |
| **RQ3:** Processing directions are followed and map clearly represents final analytical dataset. | 5 |

## Instructions

Create a new .Rmd document. For each of the research questions below, create a text header. Then complete the analysis in a code chunk (with a descriptive comment for each command). Remember to follow the Formatting Guide.

## RQ1: What is the distribution of moderate or worse air quality days by NC county in 2025?

In this research question, our goal is to create an analytic variable representing the percent of air quality observations per county in 2025 that were moderate or worse air quality.

### Data:

-   North Carolina county boundaries from US Census Bureau
-   2025 air quality data by station from [US EPA](https://www.epa.gov/outdoor-air-quality-data/download-daily-data)

```{r, eval = F}
library(tidyverse)
library(sf)

nc_counties <- st_read("https://drive.google.com/uc?export=download&id=1g9sGIikgOEubqoj97fUVoCYAKlBDVX5a")

aq <- read_csv("https://drive.google.com/uc?export=download&id=1sP4D1SdDorB6c_MZo7H4mpkHoKOaNqIT")

```

### Processing Steps:

-   In the `aq` object, each row represents a single day in 2025 at a single station. Aggregate the `aq` object to the station level (this will summarize values at that station over the year). The aggregated dataset should include two calculated columns `total_obs` (total number of observations in the year) `total_less` (total observations less than good AQI (AQI \> 50). The `group_by` command must include latitude and longitude as well as station ID.
-   Spatialize the station data using the `st_as_sf()` command. The CRS = 4326
-   Reproject the `nc_counties` and `aq` objects into CRS = 2264 (North Carolina State Plane, ft)
-   Execute a spatial join between `aq` and `nc_counties`. Drop geometry and aggregate the `aq` object to the county level (take the sum of the `total_obs` and `total_less` columns per county).
-   Calculate a `perc_less` variable representing the percent of county observations that are less than good AQI.
-   Use a table join to add your county-aggregated data to the `nc_counties` object.
-   Create a simplified dataset that includes only the following fields: `GEOID`, `NAME`, `total_obs`, `total_less`, `perc_less`.
-   Map the `perc_less` variable for North Carolina counties.

## RQ2: How does tree cover canopy vary within walking distance from bus stops in Chapel Hill, NC?

In this research question, our goal is to create an analytic variable representing the percent tree canopy cover within .5 miles of each bus stop in Chapel Hill, NC.

### Data:

-   Bus stop locations from [Chapel Hill Open Data](https://opendata-townofchapelhill.hub.arcgis.com/datasets/townofchapelhill::chapel-hill-transit-bus-stops/explore)
-   Tree canopy cover from [NLCD](https://www.mrlc.gov/data/type/nlcd-tree-canopy-cover). Each pixel value represents the percent of tree canopy cover in that pixel.

```{r, eval = F}
library(terra)
library(exactextractr)

bus_stops <- st_read("https://drive.google.com/uc?export=download&id=1jRINUl-5uBAsBcWnKTRmdUO7ZTs1EV4G")

tree_canopy <- rast("https://drive.google.com/uc?export=download&id=1_SqO-ocyLCg3g1Mv1ZbqOd5Pa1sTddps")

```

### Processing Steps:

-   Transform `bus_stops` object into EPSG:3857 to match the tree canopy projection. Note that the units in this projection are meters
-   Buffer the `bus_stops` 804.672 meters (.5 miles)
-   Use the `exact_extract()` function to add a field to the buffered bus stop object that represents the average canopy cover
-   Drop the geometry
-   Create a simplified dataset that includes only the following fields: `STOP_ID`, average tree canopy variable
-   Map the average tree canopy variable. Bus stop points (**NOT** buffers) should be visualized. You should use the `st_drop_geometry()` command and a `left_join()` to join your buffered values back to the original bus stop object

## RQ3: How accessible are bus stops to Chapel Hill addresses?

In this research question, our goal is to create an analytic variable representing the distance of each address in Chapel Hill to the nearest bus stop.

### Data:

-   Bus stop locations from [Chapel Hill Open Data](https://opendata-townofchapelhill.hub.arcgis.com/datasets/townofchapelhill::chapel-hill-transit-bus-stops/explore)
-   Chapel Hill addresses from [Chapel Hill Open Data](https://opendata-townofchapelhill.hub.arcgis.com/datasets/townofchapelhill::chapel-hill-transit-bus-stops/explore)

```{r, eval = F}
ch_addresses <- st_read("https://drive.google.com/uc?export=download&id=1fFXfEbOWjwfsT_JLeLYnaVPkbJCGYA2_")

```

### Processing Steps:

-   Transform the `ch_addresses` object into EPSG:3857 to match the bus stop object.
-   Calculate the distance (in meters) from each address to the nearest bus stop.
-   Create a simplified dataset that includes only the following fields: `OBJECTID`, `LBCSDesc`, distance variable
-   Create a map of the distance variable for Chapel Hill addresses
